---
title: "0-5"
author: "Ashley"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(nlme)
library(vegan)
library(ggeffects)
library(cowplot)
library(ggpubr)
library(sjPlot)

#Using 0-5cm depth

#New soil pH and Fe data:
ehp_soildat_deep=read.csv("fe_pH_Compiled.csv") %>% 
    mutate(plotID=as.numeric(substr(Plot, 2, 5))) %>% 
  filter(soil_depth=="0-5cm") %>% 
 arrange(plotID) %>% 
  select(-County, -Unit, -Plot, -soil_depth, -c_percent, -n_percent, -cn) 

#reading in and cleaning bacteria sequence file
 bacteria_data_1=read.csv("newdata/Bacteria/IndianaCFI16S_Prop_seqtab_0_5_2021_04_16.csv",head =TRUE, row.names=1)
bacteria_data_1<-t(bacteria_data_1) 

bacteria_data=as.data.frame(bacteria_data_1) %>% 
  rownames_to_column("id") %>% 
  separate(id, into=c("plot", "upperDepth", "lowerDepth", "amplicon")) %>% 
  mutate(plot=substr(plot, 2, 5)) %>% #38 plots
  select(-upperDepth, -lowerDepth, -amplicon) %>% 
  arrange(plot)
bacteria_plots=bacteria_data$plot
bacteria_data=bacteria_data %>% column_to_rownames(var="plot") 

#reading in and cleaning AM sequence file
amfungi_data_1=read.csv("newdata/AMF/IndianaCFIAMF_seqtab_prop_0_5_2022_02_24.csv",head =TRUE, row.names=1) 
amfungi_data_1<-t(amfungi_data_1)

amfungi_data=as.data.frame(amfungi_data_1) %>%  
  rownames_to_column("id") %>% 
  separate(id, into=c("plot", "upperDepth", "lowerDepth", "microbialGroup")) %>% 
    mutate(plot=as.numeric(substr(plot, 2, 5))) %>% 
  select(-upperDepth, -lowerDepth, -microbialGroup)  %>% 
  arrange(plot)
amfungi_plots=amfungi_data$plot
amfungi_data=amfungi_data %>% column_to_rownames(var="plot")

#reading in and cleaning Fungi sequence file
fungi_data_1=read.csv("newdata/Fungi/IndianaCFIITSProp_seqtab_0_5_2022_02_24_1.csv",head =TRUE, row.names=1)
fungi_data_1<-t(fungi_data_1)

fungi_data=as.data.frame(fungi_data_1) %>% 
  rownames_to_column("id") %>% 
  separate(id, into=c("plot", "upperDepth", "lowerDepth", "amplicon")) %>% 
  mutate(plot=as.numeric(substr(plot, 2, 5))) %>% 
  select(-upperDepth, -lowerDepth, -amplicon) %>% 
  arrange(plot)
fungi_plots=fungi_data$plot
fungi_data=fungi_data %>% column_to_rownames(var="plot")

#reading in and cleaning ECM fungi sequence file
ecm_data_1=read.csv("newdata/Fungi/IndianaCFIITSProp_seqtabEcto_0_5_2022_02_24.csv",head =TRUE, row.names=1)
ecm_data_1<-t(ecm_data_1)

ecm_data=as.data.frame(ecm_data_1) %>% 
  rownames_to_column("id") %>% 
  separate(id, into=c("plot", "upperDepth", "lowerDepth", "amplicon")) %>% 
  mutate(plot=as.numeric(substr(plot, 2, 5))) %>% 
  select(-upperDepth, -lowerDepth, -amplicon) %>% 
  arrange(plot)
ecm_plots=ecm_data$plot
ecm_data=ecm_data %>% column_to_rownames(var="plot")



#compiling environmental data: AM
am_clim=read.csv("newdata/AMF/Climate_0_5_02_2022.csv")
am_soil=read.csv("newdata/AMF/fe_pH_0_5_2022_02_15.csv")
am_geo=read.csv("newdata/AMF/Geo_0_5_02_2022.csv")
am_productivity=read.csv("newdata/AMF/PlantProd_0_5_02_2022.csv")
am_structure=read.csv("newdata/AMF/PlantStructural_0_5_02_2022.csv")
am_treediv=read.csv("newdata/AMF/PlantTreeDiv_0_5_02_2022.csv")

am_environmental_data=am_clim %>% 
  left_join(am_geo, by="X") %>% 
  left_join(am_soil, by=c("X"="Name")) %>% 
  left_join(am_productivity, by="X") %>% 
  left_join(am_structure, by="X") %>% 
  left_join(am_treediv, by="X") %>% 
  drop_na() %>% 
  separate(X, into=c("plotID", "upperDepth", "lowerDepth", "amplicon")) %>% 
  select(-"upperDepth", -"lowerDepth", -"amplicon")

#compiling environmental data: Fungi (ECM and total)
fungi_clim=read.csv("newdata/Fungi/Climate_0_5_02_2022.csv")
fungi_soil=read.csv("newdata/Fungi/fe_pH_0_5_2022_02_15.csv")
fungi_geo=read.csv("newdata/Fungi/Geo_0_5_02_2022.csv")
fungi_productivity=read.csv("newdata/Fungi/PlantProd_0_5_02_2022.csv")
fungi_structure=read.csv("newdata/Fungi/PlantStructural_0_5_02_2022.csv")
fungi_treediv=read.csv("newdata/Fungi/PlantTreeDiv_0_5_02_2022.csv")

fungi_environmental_data=fungi_clim %>% 
  left_join(fungi_geo, by="X") %>% 
  left_join(fungi_soil, by=c("X"="Name")) %>% 
  left_join(fungi_productivity, by="X") %>% 
  left_join(fungi_structure, by="X") %>% 
  left_join(fungi_treediv, by="X") %>% 
  drop_na()%>% 
  separate(X, into=c("plotID", "upperDepth", "lowerDepth", "amplicon")) %>% 
  select(-"upperDepth", -"lowerDepth", -"amplicon")

#compiling environmental data: Bacteria
bacteria_clim=read.csv("newdata/Bacteria/Climate_0_5_02_2022.csv")
bacteria_soil=read.csv("newdata/Bacteria/fe_pH_0_5_2022_02_15.csv")
bacteria_geo=read.csv("newdata/Bacteria/Geo_0_5_02_2022.csv")
bacteria_productivity=read.csv("newdata/Bacteria/PlantProd_0_5_02_2022.csv")
bacteria_structure=read.csv("newdata/Bacteria/PlantStructural_0_5_02_2022.csv")
bacteria_treediv=read.csv("newdata/Bacteria/PlantTreeDiv_0_5_02_2022.csv")

bacteria_environmental_data=bacteria_clim %>% 
  left_join(bacteria_geo, by="X") %>% 
  left_join(bacteria_soil, by=c("X"="Name")) %>% 
  left_join(bacteria_productivity, by="X") %>% 
  left_join(bacteria_structure, by="X") %>% 
  left_join(bacteria_treediv, by="X") %>% 
  drop_na()%>% 
  separate(X, into=c("plotID", "upperDepth", "lowerDepth", "amplicon")) %>% 
  select(-"upperDepth", -"lowerDepth", -"amplicon")


```


```{r AM_0-5cm_ad, warning=F, echo=F}
AM_shannon <- diversity(amfungi_data, index="shannon") 
AM_shannon=as.data.frame(AM_shannon) %>% 
  rownames_to_column("plotID") 

AM_simpson <- diversity(amfungi_data, index="simpson")
AM_simpson=as.data.frame(AM_simpson) %>% 
  rownames_to_column("plotID") 

AM_invsimpson <- diversity(amfungi_data, index="invsimpson")
AM_invsimpson=as.data.frame(AM_invsimpson) %>% 
  rownames_to_column("plotID") 

AM_species_richness = amfungi_data %>% 
  rownames_to_column("plotID") %>% 
mutate(AM_species_richness= rowSums(amfungi_data != 0)) %>% 
  select(plotID, AM_species_richness)

```

```{r Fungi_0-5cm_ad, warning=F, echo=F}
Fungi_shannon <- diversity(fungi_data, index="shannon") 
Fungi_shannon=as.data.frame(Fungi_shannon) %>% 
  rownames_to_column("plotID") 

Fungi_simpson <- diversity(fungi_data, index="simpson")
Fungi_simpson=as.data.frame(Fungi_simpson) %>% 
  rownames_to_column("plotID") 

Fungi_invsimpson <- diversity(fungi_data, index="invsimpson")
Fungi_invsimpson=as.data.frame(Fungi_invsimpson) %>% 
  rownames_to_column("plotID") 

Fungi_species_richness = fungi_data %>% 
  rownames_to_column("plotID") %>% 
mutate(Fungi_species_richness= rowSums(fungi_data != 0)) %>% 
  select(plotID, Fungi_species_richness)

```

```{r ECM_0-5cm_ad, warning=F, echo=F}
ECM_shannon <- diversity(ecm_data, index="shannon") 
ECM_shannon=as.data.frame(ECM_shannon) %>% 
  rownames_to_column("plotID") 

ECM_simpson <- diversity(ecm_data, index="simpson")
ECM_simpson=as.data.frame(ECM_simpson) %>% 
  rownames_to_column("plotID") 

ECM_invsimpson <- diversity(ecm_data, index="invsimpson")
ECM_invsimpson=as.data.frame(ECM_invsimpson) %>% 
  rownames_to_column("plotID") %>% 
  filter( ECM_invsimpson != Inf)

ECM_species_richness = ecm_data %>% 
  rownames_to_column("plotID") %>% 
mutate(ECM_species_richness= rowSums(ecm_data != 0)) %>% 
  select(plotID, ECM_species_richness)

```

```{r Bact_0-5cm_ad, warning=F, echo=F}
Bact_shannon <- diversity(bacteria_data, index="shannon") 
Bact_shannon=as.data.frame(Bact_shannon) %>% 
  rownames_to_column("plotID") 

Bact_simpson <- diversity(bacteria_data, index="simpson")
Bact_simpson=as.data.frame(Bact_simpson) %>% 
  rownames_to_column("plotID") 

Bact_invsimpson <- diversity(bacteria_data, index="invsimpson")
Bact_invsimpson=as.data.frame(Bact_invsimpson) %>% 
  rownames_to_column("plotID") 

Bact_species_richness = bacteria_data %>% 
  rownames_to_column("plotID") %>% 
mutate(Bact_species_richness= rowSums(bacteria_data != 0)) %>% 
  select(plotID, Bact_species_richness)
```

```{r diversity_metrics, echo=F, message=F, warning=F}

AM_diversity_data=am_environmental_data %>% 
  left_join(AM_shannon, by="plotID") %>%
  left_join(AM_simpson, by="plotID") %>%
  left_join(AM_invsimpson, by="plotID") %>%
  left_join(AM_species_richness, by="plotID") %>% 
  mutate(AM_species_evenness=AM_shannon/log(AM_species_richness))

Fungi_diversity_data=fungi_environmental_data %>% 
  left_join(Fungi_shannon, by="plotID") %>% 
  left_join(Fungi_simpson, by="plotID") %>%
  left_join(Fungi_invsimpson, by="plotID") %>%
   left_join(Fungi_species_richness, by="plotID") %>% 
  mutate(Fungi_species_evenness=Fungi_shannon/log(Fungi_species_richness)) 

ECM_diversity_data=fungi_environmental_data %>% 
  left_join(ECM_shannon, by="plotID") %>% 
  left_join(ECM_simpson, by="plotID") %>%
  left_join(ECM_invsimpson, by="plotID") %>%
   left_join(ECM_species_richness, by="plotID") %>% 
  mutate(ECM_species_evenness=ECM_shannon/log(ECM_species_richness)) 

Bacteria_diversity_data=bacteria_environmental_data%>% 
  left_join(Bact_shannon, by="plotID") %>%
  left_join(Bact_simpson, by="plotID") %>%
  left_join(Bact_invsimpson, by="plotID") %>%  
  left_join(Bact_species_richness, by="plotID") %>% 
  mutate(Bact_species_evenness=Bact_shannon/log(Bact_species_richness)) 

```
## What influences alpha diversity of AM fungi, ECM fungi, total fungal community,  and bacteria?


```{r shannon, message=F, warning=F, include=F, eval=F}

#what influences a-diversity of AM fungi?
#pH, VAI, bedrock depth:
am_shan=lm(AM_shannon~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=diversity_data, na.action=na.omit)
summary(am_shan)


#what influences a-diversity of total fungi?
#pH, AM dominance, BAI:
fun_shan=lm(Fungi_shannon~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=diversity_data, na.action=na.omit)
summary(fun_shan)


#what influences a-diversity of bacteria?
bact_shan=lm(Bact_shannon~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=diversity_data, na.action=na.omit)
summary(bact_shan)

tab_model(am_shan, fun_shan, bact_shan, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)
```


```{r simpsons, eval=F, include=F}
#what influences a-diversity of AM fungi?
am_sim=lm(AM_simpson~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=diversity_data, na.action=na.omit)
summary(am_sim)

#what influences a-diversity of total fungi?
#AM dominance (marginal):
fun_sim=lm(Fungi_simpson~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=diversity_data, na.action=na.omit)
summary(fun_sim)


#what influences a-diversity of bacteria?
bact_sim=lm(Bact_simpson~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=diversity_data, na.action=na.omit)
summary(bact_sim)

tab_model(am_sim, fun_sim, bact_sim, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)
```

### Analyses use inverse Simpson's Index and models were constructed for 5 groups of potential drivers: Tree species richness, productivity, and aboveground structure, soil characteristics, and climate.

## Tree species richness: 

```{r plantRichness, echo=F, message=F, warning=F}
#what influences a-diversity of AM fungi?

#plant richness metrics?
am_invsim_plantrichness=lm(AM_invsimpson~tree_richness+ AMdominance+ AMtreerichness+EMtreerichness , data=AM_diversity_data, na.action=na.omit)
#summary(am_invsim_plantrichness)
#no for 5-15cm, YES for 0-5: AM and ECM richness (neg?), total richness (pos?), not am dominance

ecm_invsim_plantrichness=lm(ECM_invsimpson~tree_richness+ AMdominance+ AMtreerichness+EMtreerichness , data=ECM_diversity_data, na.action=na.omit)
#summary(ecm_invsim_plantrichness)
#nothing with either depth 

fungi_invsim_plantrichness=lm(Fungi_invsimpson~tree_richness+ AMdominance+ AMtreerichness+EMtreerichness , data=Fungi_diversity_data, na.action=na.omit)
#summary(fungi_invsim_plantrichness)
#AM dominance (strongly positive) in 5-15, nothing in 0-5
#cor.test(Fungi_diversity_data$AMdominance,Fungi_diversity_data$Fungi_invsimpson, method="pearson")


bacteria_invsim_plantrichness=lm(Bact_invsimpson~tree_richness+ AMdominance+ AMtreerichness+EMtreerichness , data=Bacteria_diversity_data, na.action=na.omit)
#summary(bacteria_invsim_plantrichness)
#AM dominance (positive in 5-15, marginal in 0-5)
#cor.test(Bacteria_diversity_data$AMdominance,Bacteria_diversity_data$Bact_invsimpson, method="pearson")


tab_model(am_invsim_plantrichness, ecm_invsim_plantrichness, fungi_invsim_plantrichness, bacteria_invsim_plantrichness, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)



```
 

## Forest productivity: 

```{r plantProductivity, echo=F, message=F, warning=F}



#plant productivity metrics
am_invsim_plantprod=lm(AM_invsimpson~stand_age_years+ BAtotal_2020+ BAI, data=AM_diversity_data, na.action=na.omit)
#summary(am_invsim_plantprod)
#marginal effect of stand age (negative effect) in 5-10, same effect but significant in 0-5. 




ecm_invsim_plantprod=lm(ECM_invsimpson~stand_age_years+ BAtotal_2020+ BAI, data=ECM_diversity_data, na.action=na.omit)
#summary(ecm_invsim_plantprod)
#nothing in 5-10, positive effect of stand age in 0-5



fungi_invsim_plantprod=lm(Fungi_invsimpson~stand_age_years+ BAtotal_2020+ BAI, data=Fungi_diversity_data, na.action=na.omit)
#summary(fungi_invsim_plantprod)
#stand age (strong negative) in 5-10, nothing in 0-5
cor.test(Fungi_diversity_data$stand_age_years,Fungi_diversity_data$Fungi_invsimpson, method="pearson")


bacteria_invsim_plantprod=lm(Bact_invsimpson~stand_age_years+ BAtotal_2020+ BAI, data=Bacteria_diversity_data, na.action=na.omit)
#summary(bacteria_invsim_plantprod)
#stand age (strong negative) in both depths
cor.test(Bacteria_diversity_data$stand_age_years,Bacteria_diversity_data$Bact_invsimpson, method="pearson")


tab_model(am_invsim_plantprod,ecm_invsim_plantprod, fungi_invsim_plantprod, bacteria_invsim_plantprod, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)


```
 


## Canopy structure

```{r plantStructure, echo=F, message=F, warning=F}


#plant structural metrics
am_invsim_plantstructure=lm(AM_invsimpson~vert.sd+ VAI+ VCI, data=AM_diversity_data, na.action=na.omit)
#summary(am_invsim_plantstructure)
#nope


ecm_invsim_plantstructure=lm(ECM_invsimpson~vert.sd+ VAI+ VCI, data=ECM_diversity_data, na.action=na.omit)
#summary(ecm_invsim_plantstructure)
#nothing in 5-10, VAI positive with ECM diversity in 0-5 (might be connected to stand age?)

fungi_invsim_plantstructure=lm(Fungi_invsimpson~vert.sd+ VAI+ VCI, data=Fungi_diversity_data, na.action=na.omit)
#summary(fungi_invsim_plantstructure)
#nope


bacteria_invsim_plantstructure=lm(Bact_invsimpson~vert.sd+ VAI+ VCI, data=Bacteria_diversity_data, na.action=na.omit)
#summary(bacteria_invsim_plantstructure)
#nope


tab_model(am_invsim_plantstructure, ecm_invsim_plantstructure, fungi_invsim_plantstructure, bacteria_invsim_plantstructure, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)

```
 

## Soil:

```{r soil, echo=F, message=F, warning=F}

#soil 
am_invsim_soil=lm(AM_invsimpson~cn+pH+Fe_percent, data=AM_diversity_data, na.action=na.omit)
#summary(am_invsim_soil)
#summary(lm(AM_invsimpson~pH, data=AM_diversity_data))
cor.test(AM_diversity_data$pH,AM_diversity_data$AM_invsimpson, method="pearson")
#pH (positive) both depths


ecm_invsim_soil=lm(ECM_invsimpson~cn+pH+Fe_percent, data=ECM_diversity_data, na.action=na.omit)
#summary(ecm_invsim_soil)
summary(lm(ECM_invsimpson~pH, data=ECM_diversity_data))
cor.test(ECM_diversity_data$pH,ECM_diversity_data$ECM_invsimpson, method="pearson")
#nothing both depths

fungi_invsim_soil=lm(Fungi_invsimpson~cn+pH+Fe_percent, data=Fungi_diversity_data, na.action=na.omit)
#summary(fungi_invsim_soil)
#summary(lm(Fungi_invsimpson~pH, data=Fungi_diversity_data))
cor.test(Fungi_diversity_data$pH,Fungi_diversity_data$Fungi_invsimpson, method="pearson")
#pH (strongly positive) boph depths and C:N (strongly negative) 5-10


bacteria_invsim_soil=lm(Bact_invsimpson~cn+pH+Fe_percent, data=Bacteria_diversity_data, na.action=na.omit)
#summary(bacteria_invsim_soil)
#summary(lm(Bact_invsimpson~pH, data=Bacteria_diversity_data))
cor.test(Bacteria_diversity_data$pH,Bacteria_diversity_data$Bact_invsimpson, method="pearson")
#pH (strongly positive) both depths



tab_model(am_invsim_soil, ecm_invsim_soil, fungi_invsim_soil, bacteria_invsim_soil, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)


```
 

## Climate:

```{r climate, echo=F, message=F, warning=F}

#climate
am_invsim_clim=lm(AM_invsimpson~MAT+ TAP, data=AM_diversity_data, na.action=na.omit)
#summary(am_invsim_clim)
#no, both depths

ecm_invsim_clim=lm(ECM_invsimpson~MAT+ TAP, data=ECM_diversity_data, na.action=na.omit)
#summary(ecm_invsim_clim)
#nothing, both depths

fungi_invsim_clim=lm(Fungi_invsimpson~MAT+ TAP, data=Fungi_diversity_data, na.action=na.omit)
#summary(fungi_invsim_clim)
#MAT (positive) both depths

bacteria_invsim_clim=lm(Bact_invsimpson~MAT+ TAP, data=Bacteria_diversity_data, na.action=na.omit)
#summary(bacteria_invsim_clim)
#none, both depths


tab_model(am_invsim_clim, ecm_invsim_clim, fungi_invsim_clim, bacteria_invsim_clim, show.se = TRUE, show.ci = FALSE, show.std = "std2", digits = 3, digits.re = 3)

```
 

```{r species_richness, eval=F, include=F}
#what influences species richness of AM fungi?
#VAI, pH, average bedrock depth:
am_sr=lm(AM_species_richness~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=AM_diversity_data, na.action=na.omit)
summary(am_sr)


#what influences species ricness of total fungi?
#stand age, tree species richness, AM dominance, vert.sd, MAT, pH, C:N:
fun_sr=lm(Fungi_species_richness~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=Fungi_diversity_data, na.action=na.omit)
summary(fun_sr)


#what influences species richness of bacteria?
#pH:
bact_sr=lm(Bact_species_richness~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=BActeria_diversity_data, na.action=na.omit)
summary(bact_sr)



```
```{r species_evenness, eval=F, include=F}
#what influences species evenness of AM fungi?
#VAI, pH, average bedrock depth:
am_ev=lm(AM_species_evenness~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=AM_diversity_data, na.action=na.omit)
summary(am_ev)


#what influences species evenness of total fungi?
#stand age, tree species richness, AM dominance, vert.sd, MAT, pH, C:N:
fun_ev=lm(Fungi_species_evenness~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=Bacteria_diversity_data, na.action=na.omit)
summary(fun_ev)


#what influences species evenness of bacteria?
#pH:
bact_ev=lm(Bact_species_evenness~stand_age_years+ BAtotal_2020+ BAI+ tree_richness+ AMdominance+ VAI+ VCI+ vert.sd+ MAT+ TAP+ C_N+pH+Fe_percent+ avg_brockdepmi, data=Bacteria_diversity_data, na.action=na.omit)
summary(bact_ev)



```

## pH influences alpha diversity of AM fungi, bacteria, and total fungi:

```{r pH_fig, echo=F, message=F, warning=F}

#calculating the marginal effect of pH on microbial group alpha diversity 

AM_E1 = ggeffect(am_invsim_soil, terms = c("pH"), type = "random")
Fun_E1= ggeffect(fungi_invsim_soil, terms = c("pH"), type = "random")
Bact_E1= ggeffect(bacteria_invsim_soil, terms = c("pH"), type = "random")
ECM_E1= ggeffect(ecm_invsim_soil, terms=c("pH"), type="random")

a <- ggplot() +
  geom_point(data = AM_diversity_data, aes(x = pH, y = AM_invsimpson), size=3) +
  geom_line(data = AM_E1, aes(x = x, y = predicted), size = 1.5) +
  labs(x= "Soil pH", y="Alpha diversity of AM fungi" )+
  theme_cowplot()+
  geom_text(aes(x=6.9, y=10, label="r = 0.42\np = 0.04"))


b <- ggplot() +
  geom_point(data = Fungi_diversity_data, aes(x = pH, y = Fungi_invsimpson), size=3) +
  geom_line(data = Fun_E1, aes(x = x, y = predicted), size = 1.5) +
  labs(x= "Soil pH", y="Alpha diversity of Total fungi" )+
  theme_cowplot()+
  geom_text(aes(x=6.9, y=40, label="r = 0.47\np < 0.001"))


  c <- ggplot()+
  geom_point(data = Bacteria_diversity_data, aes(x = pH, y = Bact_invsimpson), size=3) +
  geom_line(data = Bact_E1, aes(x = x, y = predicted), size = 1.5) +
labs(x= "Soil pH", y="Alpha diversity of Bacteria" )+
  theme_cowplot()+
  geom_text(aes(x=6.9, y=330, label="r = 0.56\np < 0.001"))


   d <- ggplot()+
  geom_point(data = ECM_diversity_data, aes(x = pH, y = ECM_invsimpson), size=3) +
  geom_line(data = ECM_E1, aes(x = x, y = predicted), size = 1.5, linetype="dashed") +
labs(x= "Soil pH", y="Alpha diversity of ECM fungi" )+
  theme_cowplot()+
  geom_text(aes(x=6.9, y=8, label="r = -0.18\np = 0.143"))



  ggarrange(c,b,a,d, labels=c("a", "b", "c", "d" ), nrow=2, ncol=2)
  
  
  #compare with figure without marginal effects:

#   e <- ggplot(data = AM_diversity_data, aes(x = pH, y = AM_invsimpson)) +
#   geom_point( size=3) +
#   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
#   labs(x= "pH", y="Alpha diversity of AM Fungi" )+
#   theme_cowplot()+
#   geom_text(aes(x=6.9, y=10, label="r = 0.42\np = 0.04"))
# 
# 
# f <- ggplot(data = Fungi_diversity_data, aes(x = pH, y = Fungi_invsimpson)) +
#   geom_point( size=3) +
#   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
#   labs(x= "pH", y="Alpha diversity of Fungi" )+
#   theme_cowplot()+
#   geom_text(aes(x=6.9, y=40, label="r = 0.47\np < 0.001"))
# 
# 
#   g <- ggplot(data = Bacteria_diversity_data, aes(x = pH, y = Bact_invsimpson)) +
#   geom_point( size=3) +
#   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
#   labs(x= "pH", y="Alpha diversity of Bacteria" )+
#   theme_cowplot()+
#    geom_text(aes(x=6.9, y=330, label="r = 0.56\np < 0.001"))
# 
#   
#     h <- ggplot(data = ECM_diversity_data, aes(x = pH, y = Bact_invsimpson)) +
#    geom_point( size=3) +
#    geom_smooth(method="lm", size = 1.5, se=F, color="black") +
#    labs(x= "pH", y="Alpha diversity of Bacteria" )+
#    theme_cowplot()+
#    geom_text(aes(x=6.9, y=8, label="r = -0.18\np = 0.143"))
#     
#    ggarrange(e,f,g,h, labels=c("a", "b", "c", "d" ), nrow=2, ncol=2)
  
  ggsave("Figure_2.pdf", width = 10 , height = 8, units = c("in"))
```

```{r stand_age_AM_dom_fig, echo=F, message=F, warning=F}
 Bact_E2 = ggeffect(bacteria_invsim_plantprod, terms = c("stand_age_years"), type = "random")
 Fun_E2= ggeffect(fungi_invsim_plantprod, terms = c("stand_age_years"), type = "random")
 Bact_E3= ggeffect(bacteria_invsim_plantrichness, terms = c("AMdominance"), type = "random")
 Fun_E3= ggeffect(fungi_invsim_plantrichness, terms=c("AMdominance"), type="random")

 a <- ggplot() +
  geom_point(data = Bacteria_diversity_data, aes(x = stand_age_years, y = Bact_invsimpson), size=3) +
  geom_line(data = Bact_E2, aes(x = x, y = predicted), size = 1.5) +
  labs(x= "Stand age (years)", y="Alpha diversity of Bacteria" )+
  theme_cowplot()+
  geom_text(aes(x=20, y=550, label="r = -0.47\np = 0.009"))


b <- ggplot() +
  geom_point(data = Fungi_diversity_data, aes(x = stand_age_years, y = Fungi_invsimpson), size=3) +
  geom_line(data = Fun_E2, aes(x = x, y = predicted), size = 1.5) +
  labs(x= "Stand age (years)", y="Alpha diversity of Total fungi" )+
  theme_cowplot()+
  geom_text(aes(x=20, y=72, label="r = -0.72\np < 0.001"))


  c <- ggplot()+
  geom_point(data = Bacteria_diversity_data, aes(x = AMdominance, y = Bact_invsimpson), size=3) +
  geom_line(data = Bact_E3, aes(x = x, y = predicted), size = 1.5) +
labs(x= "Proportion of tree basal area\nfrom AM-associated species", y="Alpha diversity of Bacteria" )+
  theme_cowplot()+
  geom_text(aes(x=0.05, y=550, label="r = 0.46\np = 0.02"))


   d <- ggplot()+
  geom_point(data = Fungi_diversity_data, aes(x = AMdominance, y = Fungi_invsimpson), size=3) +
  geom_line(data = Fun_E3, aes(x = x, y = predicted), size = 1.5) +
labs(x= "Proportion of tree basal area\nfrom AM-associated species", y="Alpha diversity of Total fungi" )+
  theme_cowplot()+
  geom_text(aes(x=0.05, y=72, label="r = 0.65\np = 0.010"))
 
 
 
 # a <- ggplot(data = Bacteria_diversity_data, aes(x = stand_age_years, y = Bact_invsimpson)) +
 #   geom_point( size=3) +
 #   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
 #   labs(x= "Stand Age (years)", y="Alpha diversity of Bacteria" )+
 #   theme_cowplot()+
 #   geom_text(aes(x=20, y=550, label="p = 0.143\nr = -0.18"))
 # 
 # b <- ggplot(data = Fungi_diversity_data, aes(x = stand_age_years, y = Fungi_invsimpson)) +
 #   geom_point( size=3) +
 #   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
 #   labs(x= "Stand Age (years)", y="Alpha diversity of Total fungi" )+
 #   theme_cowplot()+
 #   geom_text(aes(x=20, y=72, label="p = 0.143\nr = -0.18"))
 # 
 #  c <- ggplot(data = Bacteria_diversity_data, aes(x = AMdominance, y = Bact_invsimpson)) +
 #   geom_point( size=3) +
 #   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
 #   labs(x= "Proportion of tree basal area\nfrom AM-associated species", y="Alpha diversity of Bacteria" )+
 #   theme_cowplot()+
 #   geom_text(aes(x=0.05, y=550, label="p = 0.143\nr = -0.18"))
 #  
 #   d <- ggplot(data = Fungi_diversity_data, aes(x = AMdominance, y = Fungi_invsimpson)) +
 #   geom_point( size=3) +
 #   geom_smooth(method="lm", size = 1.5, se=F, color="black") +
 #   labs(x= "Proportion of tree basal area\nfrom AM-associated species", y="Alpha diversity of Total fungi" )+
 #   theme_cowplot()+
 #   geom_text(aes(x=0.05, y=72, label="p = 0.143\nr = -0.18"))
   
   ggarrange(a,b,c,d, labels=c("a", "b", "c", "d"), nrow=2, ncol=2)
   
     ggsave("Figure_3marginal.pdf", width = 10 , height = 8, units = c("in"))

```

```{r correlations, echo=F, message=F, warning=F}

library(dichromat)
cor_df=fungi_environmental_data %>% 
  dplyr::select(-plotID, -GPS_LAT, -GPS_LON, -TAP, -n_percent, -c_percent, -Fe_percent, -BAtotal_2020, -BAI, -VAI, -VCI, -vert.sd, -tree_richness, -AMtreerichness, -EMtreerichness) %>% 
  drop_na()
  
my_pal=colorRampPalette(c("blue","white","red"))

pdf(file = "CFI_alpha_correlations.pdf")
corrplot::corrplot(cor(cor_df), method="circle", type="upper", diag=F, addCoef.col = "black", col = my_pal(n=200))
dev.off()

#-plotID, -GPS_LAT, -GPS_LON, -TAP, -n_percent, -c_percent, -Fe_percent, -BAtotal_2020, -BAI, -VAI, -VCI, -vert.sd, -tree_richness, -AMtreerichness, -EMtreerichness
```

